\textls[-0]{\thetitle}

### 2023-03-14 canonicitymodel.tex
Working on canonicity model.

Like most logical relation proof,  we interpret each judgement and inductively interpret each syntax piece. We are working on an intrinsic setting thus even if we omit context for brevity, the syntax piece is still well-typed. 
% We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model). Interested audience can refer to \citet{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} for a complete technical treatements.

\begin{align*}
  \denotesC{\goodCtx{\Gm}{k}} & \text{ is a Set of closed substitution} \goodSub{\cdot}{\_}{\Gm} \\
  % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
  \denotesC{\goodType{\Gm}{T}{j}} & \text{ is a function maps to $\Set{j}$ s.t.} \\
  & \text{for arbitrary } \gm \in \denotesC{\goodCtx{\Gm}{}}, \denotesC{\goodType{\Gm}{T}{}}(\gm) \text{ is a an element of } \Set{j}  \\ 
  % &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
  \denotesC{\goodSub{\Gm}{\gm}{\Delta}} & \text{ is a function}: \denotesC{\goodCtx{\Gm}{i}} \to \denotesC{\goodCtx{\Delta}{i}} \text{ s.t. } \denotesC{\goodSub{\Gm}{\gm}{\Delta}}(\sigma) \coloneqq \SubstComp{\sigma}{\gm} \\
  % &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
  \denotesC{\goodTerm{\Gm}{t}{T}} & \text{ is a dependent function}: \prod_{\gm \in \denotesC{\goodCtx{\Gm}{i}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)  \text{ s.t. } \denotesC{\goodTerm{\Gm}{t}{T}}(\gm) \coloneqq \sub{t}{\gm} \\
  \denotesC{\goodType{\Gm}{\top}{}}(\gm) &= \{t : t \equiv \unit \} \\
  \denotesC{\goodType{\Gm}{\bot}{}}(\gm) &= \emptyset \\
  \denotesC{\goodType{\Gm}{\cB}{}}(\gm) &= \{t : t \equiv \true \text{ or } t \equiv\false \} \\
  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm) &= \{t : t \equiv \eqrefl{\sub{a}{\gm}} \} \text{ when } \sub{a}{\gm} \equiv \sub{b}{\gm} \\ 
  &= \emptyset \text{ otherwise } \\ 
  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}} &= \\
  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}} &= \\
  \denotesC{\goodCtx{\cdot}{}} &= \{\SubstEmp\} \text{  the singleton set with identity substitution} \\ 
  \denotesC{\goodCtx{\Gm, T}{}} &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}, t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}  \} \\
  \denotesC{\goodTerm{\Gm}{\unit}{\top}}(\gm) &= \unit\\
  \denotesC{\goodTerm{\Gm}{\true}{\cB}}(\gm) &= \true \\
  \denotesC{\goodTerm{\Gm}{\false}{\cB}}(\gm) &= \false \\
  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm) &= \denotesC{a}(\gm) \text{ if } \denotes{c}(\gm) \equiv \true \\
  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm) &= \denotesC{b}(\gm) \text{ if } \denotes{c}(\gm) \equiv \false \\
  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm) &= \eqrefl{\sub{t}{\gm}}\\
  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\app{t}}{B}} &= \\
  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}} &= \\
  \denotesC{\goodTerm{\Gm}{\snd{t}}{T}} &= \\
  \denotesC{\goodType{\Gm}{\cU_j}{j}}(\gm) &= \{t : \goodTerm{\cdot}{t}{\El{T}}\} \quad \text{     for some } \goodTerm{\cdot}{T}{\cU_j}  \\
  \denotesC{\goodTerm{\Gm}{\codety{T}}{\cU_j}}(\gm) &= \denotesC{T}(\gm) \\ 
  \denotesC{\goodType{\Gm}{\El{T}}{j}}(\gm) &= \denotesC{T}(\gm) \\
\end{align*}

@@

>>>>

Like most logical relation proof,  we interpret each judgement and inductively interpret each syntax piece. We are working on an intrinsic setting thus even if we omit context for brevity, the syntax piece is still well-typed. However, unlike most logical relation proof, we are having a proof-relevant logical realtion, essential for modeling universes properly~\cite{coquand2018canonicity}. 
% We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model). Interested audience can refer to \citet{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} for a complete technical treatements.

\begin{align*}
  \denotesC{\goodCtx{\Gm}{k}} &= \{\gm : \goodSub{\cdot}{\gm}{\Gm} \} \to \Set{k} \quad \text{ i.e., sets indexed by closed substitution}\\
  % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
  \denotesC{\goodType{\Gm}{T}{j}} & \text{ is a function maps to Set, s.t. for arbitrary } \goodSub{\cdot}{\gm}{\Gm} , \\
  & \text{ and arbitrary } \goodTerm{\cdot}{t}{\sub{T}{\gm}}, \quad \text{ we have } \denotesC{\goodType{\Gm}{T}{}}(\gm)(t) \in \Set{j}  \\ 
  % &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
  \denotesC{\goodSub{\Gm}{\gm}{\Delta}} & \text{ is a function}: \denotesC{\goodCtx{\Gm}{i}} \to \denotesC{\goodCtx{\Delta}{i}} \text{ s.t. } \denotesC{\goodSub{\Gm}{\gm}{\Delta}}(\sigma) \coloneqq \SubstComp{\sigma}{\gm} \\
  % &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
  \denotesC{\goodTerm{\Gm}{t}{T}} & \text{ is a dependent function}: \prod_{\gm \in \denotesC{\goodCtx{\Gm}{i}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)  \text{ s.t. } \denotesC{\goodTerm{\Gm}{t}{T}}(\gm) \coloneqq \sub{t}{\gm} \\
  \denotesC{\goodType{\Gm}{\top}{}}(\gm)(t) &= \{\star \ | \text{  when } t \equiv \unit \} \\
  \denotesC{\goodType{\Gm}{\bot}{}}(\gm) &= \emptyset \\
  \denotesC{\goodType{\Gm}{\cB}{}}(\gm) &= \{t : t \equiv \true \text{ or } t \equiv\false \} \\
  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm) &= \{t : t \equiv \eqrefl{\sub{a}{\gm}} \} \text{ when } \sub{a}{\gm} \equiv \sub{b}{\gm} \\ 
  &= \emptyset \text{ otherwise } \\ 
  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}} &= \\
  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}} &= \\
  \denotesC{\goodCtx{\cdot}{}} &= \{\SubstEmp\} \text{  the singleton set with identity substitution} \\ 
  \denotesC{\goodCtx{\Gm, T}{}} &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}, t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}  \} \\
  \denotesC{\goodTerm{\Gm}{\unit}{\top}}(\gm) &= \unit\\
  \denotesC{\goodTerm{\Gm}{\true}{\cB}}(\gm) &= \true \\
  \denotesC{\goodTerm{\Gm}{\false}{\cB}}(\gm) &= \false \\
  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm) &= \denotesC{a}(\gm) \text{ if } \denotes{c}(\gm) \equiv \true \\
  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm) &= \denotesC{b}(\gm) \text{ if } \denotes{c}(\gm) \equiv \false \\
  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm) &= \eqrefl{\sub{t}{\gm}}\\
  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\app{t}}{B}} &= \\
  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}} &= \\
  \denotesC{\goodTerm{\Gm}{\snd{t}}{T}} &= \\
  \denotesC{\goodType{\Gm}{\cU_j}{j}}(\gm) &= \{t : \goodTerm{\cdot}{t}{\El{T}}\} \quad \text{     for some } \goodTerm{\cdot}{T}{\cU_j}  \\
  \denotesC{\goodTerm{\Gm}{\codety{T}}{\cU_j}}(\gm) &= \denotesC{T}(\gm) \\ 
  \denotesC{\goodType{\Gm}{\El{T}}{j}}(\gm) &= \denotesC{T}(\gm) \\
\end{align*}

@@

The counter-intuitive part of the proof-relevant logical relation is the treatment of proposition---a proposition is encoded as a subsingleton data. For example, $\denotesC{\goodType{\Gm}{\top}{}} = $ $ \{\star \ | \text{  when } t \equiv \unit \}$ represents a singleton set when $t \equiv \unit$ or an empty set otherwise. However, not all the denotation of types are subsingleton---the salient counter-examples are $\denotes{\goodType{\Gm}{\cU}{}}$ and $\denotes{\goodType{\Gm}{\TyPi{A}{B}}{}}$. The non-trivial interpretation on these types is the main reason we use proof-relevant logical relations.
